---
title: "<b>Most Read Authors</b>"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plotly)
library(dplyr)
library(tidyr)
library(lubridate)
library(janitor)
library(ggplot2)
```

## Data Import
```{r books, echo=FALSE}
#Read CSV file 
books <- read.csv("books.csv")
```

### *Summarizing Read Counts by Author*
```{r most read authors}
most_read_authors <- books %>%
  group_by(authors) %>%
  summarize(
    total_ratings = sum(work_ratings_count, na.rm = TRUE), 
    num_books = n() #Counting the number of books per author
  ) %>%
  arrange(desc(total_ratings))

head(most_read_authors, 50)
```


### *Displaying  Top 10 Most Read Authors*
```{r Dislaying Top 10 Most Read Authors}
ggplot(head(most_read_authors, 10), aes(x = total_ratings, y = reorder(authors, total_ratings))) +
  geom_bar(stat = "identity", aes(fill = total_ratings), show.legend = FALSE) +
  labs(
    title = "Top 10 Most Read Authors by Rating",
    x = "Total Ratings Count",
    y = "Authors"
  ) +
  scale_fill_gradient(low = "red", high = "darkred") + 
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5), 
    axis.text.y = element_text(size = 10), 
    axis.text.x = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"), 
    panel.grid.major.y = element_blank() 
  ) +
  scale_x_continuous(labels = scales::comma)
```

**Top 10 Most Read Authors**

This bar chart, "Top 10 Most Read Authors by Rating" depicts the percentage of total ratings each author received relative to the dataset.
The top 10 most read authors are:


J.K. Rowling & Mary GrandPré: 14,126,791 ratings  
Suzanne Collins: 9,161,124 ratings  
Stephenie Meyer: 8,631,937 ratings  
Stephen King: 7,297,834 ratings  
Rick Riordan: 5,717,999 ratings  
J.R.R. Tolkien: 5,602,790 ratings  
Dan Brown: 5,332,589 ratings  
Cassandra Clare: 4,631,019 ratings  
Nicholas Sparks: 4,501,871 ratings  
John Green: 4,312,589 ratings 


