---
title: "Top Users by To Read"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(plotly)
library(readr)
library(knitr)
library(tidyverse)
ratings <- read_csv("../Data/ratings.csv")
to_read <- read_csv("../Data/to_read.csv")
books_with_tags <- read_csv("../Data/books_with_tags.csv")
```
In this section, we'll be looking at the top users on GoodReads who have the most "to-read" books. 

### Most Wanted-To-Read Books 

Let's first look at the most popular books our top 15 users want to read.
```{r, echo=FALSE}
#Count how many to-reads each user has
to_read_counts <- to_read %>%
  group_by(user_id) %>%
  summarize(total_to_read = n(), .groups = 'drop') %>%
  arrange(desc(total_to_read))

#Select the top 15 users with the most to-reads
top_user_ids_to_read <- to_read_counts %>%
  head(15)

#Create dataframe with all the books the top 15 users said they wanted to read. Count how many times these books were selected by users. Select the top 15 to-read books by the top 15 users
top_to_read <- semi_join(to_read, top_user_ids_to_read, by = c("user_id"))

top_to_read_count <- top_to_read %>%
  count(book_id) %>%
  arrange(desc(n)) %>%
  head(15) 

#Add book information to the 15 books
top_ratings_titles <- left_join(top_to_read_count, books_with_tags, by = c("book_id")) %>%
  select(title, authors, n) %>%
  rename("Number of Top Users" = n) %>%
  rename("Title" = title) %>%
  rename("Authors" = authors)


kable(top_ratings_titles, caption = "Top 15 To-Read books for Top 15 Users", align = "c")

```
The books that the most of the top 15 users wanted to read is "The Catcher in the Rye" by J.D. Salinger, followed by "The Fault in Our Stars" by John Green and "A Tale of Two Cities" by Charles Dickens.

### Top Genres

Now, let's look at the top genres for the 15 Users using distinct book titles. 

```{r, echo=FALSE}
top_to_read_genres <- left_join(top_to_read, books_with_tags, by = c("book_id")) %>%
  group_by(user_id) %>% 
  pivot_longer(cols = starts_with("top"),
               names_to = "tag_type",
              values_to = "tag") %>%
  group_by(user_id, tag) %>%
  summarise(tag_count = n(), .groups = "drop") %>%
  drop_na(tag) %>%
  group_by(user_id) %>%
  slice_max(order_by = tag_count, n = 3) %>%
  ungroup()

genres_count_to_read <- top_to_read_genres %>%
  count(tag) %>%
  arrange(desc(n))

genres_count_to_read %>%
plot_ly( x = ~tag, y = ~n, type = "bar") %>%
  layout(
    title = "Top Genres for Top 15 Users",
    xaxis = list(title = "Genres"),
    yaxis = list(title = "Values")
  )
```
Now let's look at the top genres of the top users' books. 

```{r, echo=FALSE}
distinct_to_read_books <- left_join(top_to_read, books_with_tags, by = c("book_id")) %>%
  pivot_longer(cols = starts_with("top"),
               names_to = "tag_type",
              values_to = "tag") %>%
  distinct(book_id, .keep_all= TRUE) %>%
  filter(!is.na(tag)) %>%
  select(-user_id) %>%
  count(tag) %>%
  filter(n >= 5) %>%
  arrange(desc(n)) 

kable(distinct_to_read_books, caption = "Top To-Read genres for Top 15 Users", align = "c")

distinct_to_read_books %>%
plot_ly( x = ~tag, y = ~n, type = "bar") %>%
  layout(
    title = "Top Genres for Top 15 To-Read Users",
    xaxis = list(title = "Genres"),
    yaxis = list(title = "Values")
  )

```

The most popular to-read genre among the top 15 users is `r distinct_to_read_books$tag[which.max(distinct_to_read_books$n)]`. The next most popular genres are `r distinct_to_read_books$tag[order(-distinct_to_read_books$n)][2]` and `r distinct_to_read_books$tag[order(-distinct_to_read_books$n)][3]`.